Dominance Regions for Imaginary $\bfg$-vectors

### 2020-06-14 dominance_regions.tex
Lemma 3.6 may only be correct when i is even

\documentclass{amsart}
\usepackage{amsmath,amssymb,latexsym,color}
\usepackage[margin=1in]{geometry}

@@

  Observe that $\phi_{t_2}=\phi_{t_{-2}}^{-1}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_{t_2}(\lambda)=\begin{cases} ((bc-1)\lambda_1+b\lambda_2, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$;}\\ (-\lambda_1, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2<0$;}\\ (-\lambda_1+b\lambda_2, -\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2\ge 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2<0$;}\end{cases}
  \end{equation}
  and $\phi_{t_2}^{-1}=\phi_{t_{-2}}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:backward two step mutation}
    \phi_{t_2}^{-1}(\lambda)=\begin{cases} (-\lambda_1-b\lambda_2, c\lambda_1+(bc-1)\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2\le 0$;}\\ (-\lambda_1-b\lambda_2, -\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2>0$;}\\ (-\lambda_1, c\lambda_1-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1\le 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1>0$;}\end{cases}
  \end{equation}

@@

  \begin{lemma}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have
    \begin{align*}
      \phi_{t_{2j}}(\lambda)&=(u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2,-u_{2j,-}\lambda_1-u_{2j-1,-}\lambda_2);\\
      \phi_{t_{2j+1}}(\lambda)&=(-u_{2j+1,+}\lambda_1-u_{2j,+}\lambda_2,u_{2j+2,-}\lambda_1+u_{2j+1,-}\lambda_2).
    \end{align*}
  \end{lemma}
  \begin{proof}
    We work by induction on $j$, the case $j=0$ being clear from the definitions.
    As observed above, we have $\phi_{t_{2j}}=\phi_{t_2}^j$.
    For $\lambda\in\cI$, the action of $\phi_{t_2}$ from \eqref{eq:forward two step mutation} can be rewritten as
    \[\phi_{t_2}(\lambda)=(u_{3,-}\lambda_1+u_{2,+}\lambda_2,-u_{2,-}\lambda_1-u_{1,+}\lambda_2),\]
    and so
    \begin{align*}
      \phi_{t_{2j+2}(\lambda)}&=\phi_{t_2}\phi_{t_{2j}}(\lambda)\\
      &=\big( (u_{3,-}u_{2j+1,+}-u_{2,+}u_{2j,-})\lambda_1+(u_{3,-}u_{2j,+}-u_{2,+}u_{2j-1,-})\lambda_2,\\
      &\qquad -(u_{2,-}u_{2j+1,+}-u_{1,+}u_{2j,-})\lambda_1-(u_{2,-}u_{2j,+}-u_{1,+}u_{2j-1,-})\lambda_2 \big)\\
      &=(u_{2j+3,+}\lambda_1+u_{2j+2,+}\lambda_2, -u_{2j+2,-}\lambda_1-u_{2j+1,-}\lambda_2)
    \end{align*}
    where the last equality uses \eqref{eq:long Chebyshev recursion}.

    Similarly, using $\phi_{t_1}(\lambda)=(-\lambda_1,c\lambda_1+\lambda_2)$ for $\lambda\in\cI$, together with the basic Chebyshev recursion and the equality $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$ gives the given formula for $\phi_{t_{2j+1}}$ from that of $\phi_{t_{2j}}$.
  \end{proof}

  The standard Chebyshev polynomials (of the second kind) are defined by the recursion $u_{i+1}=ru_i-u_{i-1}$, $u_1=1$, $u_0=0$, which can be computed explicitly as
  \[u_i(r)=\frac{1}{2^i\sqrt{r^2-4}}\left(\big(r+\sqrt{r^2-4}\big)^i-\big(r-\sqrt{r^2-4}\big)^i\right).\]
  By the equalities 
  \[u_{i,\varepsilon}=\begin{cases} \frac{\sqrt{b}}{\sqrt{c}}u_i(\sqrt{bc}) & \text{if $i$ is even and $\varepsilon=+$;}\\ \frac{\sqrt{c}}{\sqrt{b}}u_i(\sqrt{bc}) & \text{if $i$ is even and $\varepsilon=-$;}\\ u_i(\sqrt{bc}) & \text{if $i$ is odd;} \end{cases}\]
  it follows that $u_{i,\varepsilon}$ can be computed explicitly as
  \[u_{i,\varepsilon}=\begin{cases} \frac{\sqrt{b}}{2^i\sqrt{c(bc-4)}}\left(\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i\right) & \text{if $i$ is even and $\varepsilon=+$;}\\ \frac{\sqrt{c}}{2^i\sqrt{b(bc-4)}}\left(\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i\right) & \text{if $i$ is even and $\varepsilon=-$;}\\ \frac{1}{2^i\sqrt{bc-4}}\left(\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i\right) & \text{if $i$ is odd.} \end{cases}\]
  Note that $u_{2j+1,+}=u_{2j+1,-}$ for all $j\in\ZZ$.
  Moreover, since $u_{-i,\varepsilon}=-u_{i,\varepsilon}$, these formulas can also easily be seen to hold for $i<0$.

>>>>

\documentclass{amsart}
\usepackage{amsmath,amssymb,latexsym,color}
\usepackage[margin=1in]{geometry}
\usepackage{tikz}

@@

  Observe that $\phi_{t_1}$ is the piecewise-linear map given by
  \[\phi_{t_1}(\lambda)=\begin{cases} (-\lambda_1,c\lambda_1+\lambda_2) & \text{if $\lambda_1\ge0$;}\\ (-\lambda_1,\lambda_2) & \text{if $\lambda_1<0$;} \end{cases}\]
  and $\phi_{t_2}=\phi_{t_{-2}}^{-1}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_{t_2}(\lambda)=\begin{cases} ((bc-1)\lambda_1+b\lambda_2, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$;}\\ (-\lambda_1, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2<0$;}\\ (-\lambda_1+b\lambda_2, -\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2\ge 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2<0$;}\end{cases}
  \end{equation}
  while $\phi_{t_2}^{-1}=\phi_{t_{-2}}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:backward two step mutation}
    \phi_{t_2}^{-1}(\lambda)=\begin{cases} (-\lambda_1-b\lambda_2, c\lambda_1+(bc-1)\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2\le 0$;}\\ (-\lambda_1-b\lambda_2, -\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2>0$;}\\ (-\lambda_1, c\lambda_1-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1\le 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1>0$;}\end{cases}
  \end{equation}

@@

  \begin{lemma}
    \label{le:imaginary transformations}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have
    \begin{align*}
      \phi_{t_{2j}}(\lambda)&=(u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2,-u_{2j,-}\lambda_1-u_{2j-1,-}\lambda_2);\\
      \phi_{t_{2j+1}}(\lambda)&=(-u_{2j+1,+}\lambda_1-u_{2j,+}\lambda_2,u_{2j+2,-}\lambda_1+u_{2j+1,-}\lambda_2).
    \end{align*}
  \end{lemma}
  \begin{proof}
    We work by induction on $j$, the case $j=0$ being clear from the definitions.
    As observed above, we have $\phi_{t_{2j}}=\phi_{t_2}^j$.
    For $\lambda\in\cI$, the action of $\phi_{t_2}$ from \eqref{eq:forward two step mutation} can be rewritten as
    \[\phi_{t_2}(\lambda)=(u_{3,-}\lambda_1+u_{2,+}\lambda_2,-u_{2,-}\lambda_1-u_{1,+}\lambda_2),\]
    and so
    \begin{align*}
      \phi_{t_{2j+2}(\lambda)}&=\phi_{t_2}\phi_{t_{2j}}(\lambda)\\
      &=\big( (u_{3,-}u_{2j+1,+}-u_{2,+}u_{2j,-})\lambda_1+(u_{3,-}u_{2j,+}-u_{2,+}u_{2j-1,-})\lambda_2,\\
      &\qquad -(u_{2,-}u_{2j+1,+}-u_{1,+}u_{2j,-})\lambda_1-(u_{2,-}u_{2j,+}-u_{1,+}u_{2j-1,-})\lambda_2 \big)\\
      &=(u_{2j+3,+}\lambda_1+u_{2j+2,+}\lambda_2, -u_{2j+2,-}\lambda_1-u_{2j+1,-}\lambda_2)
    \end{align*}
    where the last equality uses \eqref{eq:long Chebyshev recursion}.

    Similarly, using $\phi_{t_1}(\lambda)=(-\lambda_1,c\lambda_1+\lambda_2)$ for $\lambda\in\cI$, together with the basic Chebyshev recursion and the equality $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$ gives the claimed formula for $\phi_{t_{2j+1}}$ from that of $\phi_{t_{2j}}$.
  \end{proof}

  \begin{lemma}
    For $i\ge2$ and $\lambda\in\cI$, the region $\phi_{t_i}^{-1}\cC_{t_i}(\phi_{t_i}\lambda)\subset\RR^2$ is the polytope determined by the following inequalities:
    \begin{align*}
      u_{i,-}x+u_{i-1,+}y &\le u_{i,-}\lambda_1+u_{i-1,+}\lambda_2;\\
      u_{i+1,-}x+u_{i,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2;\\
      -u_{i-1,-}x+u_{i,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2;\\
      -u_{i-1,-}x-u_{i-2,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2.
    \end{align*}

    given as in the figure below:
    \[
      \begin{tikzpicture}
        \filldraw[color=lightgray!30] (47/10,-5) --(3.5,-3) -- (0,26/10) -- (-26/4,0) -- (-16/4,-5);
        \draw[-] (-8,0) -- (5,0);
        \draw[-] (0,-5) -- (0,5);
        %\filldraw[color=lightgray!30] (0,0) -- (5,-2.113) -- (5,-5) -- (3.17,-5) -- (0,0);
        \draw[-] (0,0) -- (5,-2.113);
        \draw[-] (0,0) -- (3.17,-5);
        \node at (1,-0.9) {$\cI$};
        \draw[fill=black] (3.5,-3) circle (2pt);
        \node[below right] at (3.5,-3) {$\lambda$};
        \draw[->,thick] (3.5,-3) -- (47/10,-5);
        \draw[-,thick] (3.5,-3) -- (0,26/10);
        \draw[-,thick] (0,26/10) -- (-26/4,0);
        \draw[->,thick] (-26/4,0) -- (-16/4,-5);
      \end{tikzpicture}
    \]
  \end{lemma}
  \begin{proof}
    Following Lemma~\ref{le:imaginary transformations}, we consider even and odd sequences of mutations separately.

    For $i=2j$, $j>0$, and $\lambda\in\cI$, we get $\cC_{t_i}(\phi_{t_i}\lambda)\subset\RR^2$ is given by the inequalities 
    \[ x\le u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2 \qquad\text{and}\qquad y\le u_{2j,-}\lambda_1+u_{2j-1,-}\lambda_2. \]

    For $i=2j+1$, $j>0$, and $\lambda\in\cI$, we get $\cC_{t_i}(\phi_{t_i}\lambda)\subset\RR^2$ is given by the inequalities 
    \[ x\le u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2 \qquad\text{and}\qquad y\le u_{2j+2,-}\lambda_1+u_{2j+1,-}\lambda_2. \]
  \end{proof}

  The standard Chebyshev polynomials (of the second kind) are defined by the recursion $u_{i+1}=ru_i-u_{i-1}$, $u_1=1$, $u_0=0$, which can be computed explicitly as
  \[u_i(r)=\frac{1}{2^i\sqrt{r^2-4}}\left(\big(r+\sqrt{r^2-4}\big)^i-\big(r-\sqrt{r^2-4}\big)^i\right).\]
  By the equalities 
  \[u_{i,\varepsilon}=\begin{cases} \frac{\sqrt{b}}{\sqrt{c}}u_i(\sqrt{bc}) & \text{if $i$ is even and $\varepsilon=+$;}\\ \frac{\sqrt{c}}{\sqrt{b}}u_i(\sqrt{bc}) & \text{if $i$ is even and $\varepsilon=-$;}\\ u_i(\sqrt{bc}) & \text{if $i$ is odd;} \end{cases}\]
  it follows that $u_{i,\varepsilon}$ can be computed explicitly as
  \[u_{i,\varepsilon}=\begin{cases} \frac{\sqrt{b}}{2^i\sqrt{c(bc-4)}}\left(\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i\right) & \text{if $i$ is even and $\varepsilon=+$;}\\ \frac{\sqrt{c}}{2^i\sqrt{b(bc-4)}}\left(\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i\right) & \text{if $i$ is even and $\varepsilon=-$;}\\ \frac{1}{2^i\sqrt{bc-4}}\left(\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i\right) & \text{if $i$ is odd.} \end{cases}\]
  \begin{remark}
    Note that $u_{2j+1,+}=u_{2j+1,-}$ for all $j\in\ZZ$.
    Moreover, since $u_{-i,\varepsilon}=-u_{i,\varepsilon}$, these formulas can also easily be seen to hold for $i<0$.
  \end{remark}
